---
title: "Chicago Crime Data Visualizations"
author: "Vetrie Senthilkumar"
date: "April 26, 2018"
output: html_document
---

Load Chicago Crime Data from 2017 into R and load libraries. Convert Date variable to Date type

```{r, message=FALSE}
crime_data = read.csv("Crimes_-_2001_to_present (1).csv")

library(dplyr)
library(ggplot2)
library(stringr)
```

Quantitative EDA

```{r}
dim(crime_data)
head(crime_data)
summary(crime_data)
str(crime_data)

```

Graphs crime occurences for each of Chicago's 20 Districts

```{r}

ggplot(data = crime_data, aes(x = crime_data$District)) + 
  geom_bar() +
  labs(title = "Crime Occurences Per District", x = "Chicago Districts", y = "Number of Crime Occurences") + 
  scale_x_continuous(breaks = seq(min(crime_data$District), 25, by = 1)) 

```

Graphs 20 most common locations that crimes occur in

```{r}
crime_spots = count(crime_data, Location.Description)
crime_spots = arrange(crime_spots, desc(n))
crime_spots = head(crime_spots, 20)

crime_data_locations = subset(crime_data, crime_data$Location.Description %in% crime_spots$Location.Description)

ggplot(data = crime_data_locations, aes(x = crime_data_locations$Location.Description)) +
  geom_bar() + 
  labs(title = "Most Common Locations Crimes Occur In", x = "Types of Locations", y = "Number of Crime Occurrences") + theme(axis.text.x = element_text(angle = 90, hjust = 1))




```

Graphs most common types of crime
```{r}

crime_types = count(crime_data, Primary.Type)

ggplot(data = crime_data, aes(x = crime_data$Primary.Type))  +
  geom_bar() + 
  labs(title = "Crime Rates for Different Types of Crimes", x = "Types of Crimes", y = "Number of Occurrences") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

Grpahs fluctuations in overall crime rate during 2017
```{r}
crime_data$Date = as.character(crime_data$Date)

date_pattern = "[0-9]{1,2}/[0-9]{1,2}/[0-9]{2}"
crime_data$Date = str_extract(crime_data$Date, date_pattern)

crime_dates = count(crime_data, Date)
crime_dates$Date = as.Date.character(crime_dates$Date, format = "%m/%d/%y")

subset_arrests = subset(crime_data, Arrest == TRUE)
arrest_dates = count(subset_arrests, subset_arrests$Date)
colnames(arrest_dates) = c("Date", "n")
arrest_dates$Date = as.Date.character(arrest_dates$Date, format = "%m/%d/%y")

ggplot(crime_dates, aes(x = Date)) +
  geom_line(aes(y = crime_dates$n, color = "Crime Rate")) +
  geom_line(data = arrest_dates, aes(y = arrest_dates$n, color = "Arrest Rate")) +
  labs(title = "Crime Rate vs Arrest Rate", y = "Number of Occurences") +
  scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
  

```

Changes in crime rate by month for 5 most frequent types of crime


```{r}
most_frequent_crime_types = tail(arrange(crime_types, n), 5)
most_frequent_crime_types

assault_rate = count(subset(crime_data, crime_data$Primary.Type == "ASSAULT"), Date)
assault_rate$Date = as.Date.character(assault_rate$Date, format = "%m/%d/%y")

deceptive_pratice_rate = count(subset(crime_data, crime_data$Primary.Type == "DECEPTIVE PRACTICE"), Date)
deceptive_pratice_rate$Date = as.Date.character(deceptive_pratice_rate$Date, format = "%m/%d/%y")

criminal_damage_rate = count(subset(crime_data, crime_data$Primary.Type == "CRIMINAL DAMAGE"), Date)
criminal_damage_rate$Date = as.Date.character(criminal_damage_rate$Date, format = "%m/%d/%y")

battery_rate = count(subset(crime_data, crime_data$Primary.Type == "BATTERY"), Date)
battery_rate$Date = as.Date.character(battery_rate$Date, format = "%m/%d/%y")

theft_rate = count(subset(crime_data, crime_data$Primary.Type == "THEFT"), Date)
theft_rate$Date = as.Date.character(theft_rate$Date, format = "%m/%d/%y")

# ggplot(crime_dates, aes(x = Date)) + 
#   geom_line(data = assault_rate, aes(y = n, color = "Assault")) +
#   geom_line(data = deceptive_pratice_rate, aes(y = n, color = "Deceptive Practice")) + 
#   geom_line(data = criminal_damage_rate, aes(y = n, color = "Criminal Damage")) +
#   geom_line(data = battery_rate, aes(y = n, color = "Battery")) + 
#   geom_line(data = theft_rate, aes(y = n, color = "Theft")) + 
#   scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y") +
#   theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
#   ylim(0,25) + 
#   labs(title = "Fluctuations in Different Crime Rates", y = "Number of Occurrences")
  
ggplot(crime_dates, aes(x = Date))  +
  geom_line(data = assault_rate, aes(y = n), colour = "red") +
  scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(legend.position = "none") +
  labs(title = "Assaults Committed over Time", x = "Date", y = "Number of Assaults")

ggplot(crime_dates, aes(x = Date, color = "blue"))  +
  geom_line(data = deceptive_pratice_rate, aes(y = n), colour = "orange") +
  scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(legend.position = "none") + 
  labs(title = "Deceptive Practice Activities over Time", x = "Date", y = "Number of Deceptive Practice Incidents")

ggplot(crime_dates, aes(x = Date))  +
  geom_line(data = criminal_damage_rate, aes(y = n), colour = "green") +
  scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(legend.position = "none") +
  labs(title = "Criminal Damage Incidents over Time", x = "Date", y = "Number of Criminal Damage Incidents")

ggplot(crime_dates, aes(x = Date))  +
  geom_line(data = battery_rate, aes(y = n), colour = "blue") +
  scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(legend.position = "none") + 
  labs(title = "Batteries Commited over Time", x = "Date", y = "Number of Batteries")

ggplot(crime_dates, aes(x = Date))  +
   geom_line(data = theft_rate, aes(y = n), color = "pink") +
  scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

15 most dangerous streets with highest crime rates
```{r}

pattern = "[NWES].+"

crime_data$Street = str_extract(crime_data$Block, pattern)
dangerous_streets = count(crime_data, Street)
dangerous_streets = arrange(dangerous_streets, desc(n))

#subset(crime_data, crime_data$Street %in% head(dangerous_streets$Street,5)) 

ggplot(data = subset(crime_data, crime_data$Street %in% head(dangerous_streets$Street, 15)), aes(x = Street)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "Most Dangerous Streets", x = "Street Name", y = "Number of Crime Incidents")
  

```






